Consistent Polyhedral Surrogates via Embedding

### 2019-05-30 notes/jessie-proof-sketches.tex
Added top k ideas into jessie-proof-sketches

\newcommand{\reals}{\mathbb{R}}
\newcommand{\posreals}{\reals_{>0}}%{\reals_{++}}
\newcommand{\dom}{\mathrm{dom}}

\newcommand{\prop}[1]{\Gamma[#1]}
\newcommand{\eliccts}{\mathrm{elic}_\mathrm{cts}}
\newcommand{\eliccvx}{\mathrm{elic}_\mathrm{cvx}}
\newcommand{\elicpoly}{\mathrm{elic}_\mathrm{pcvx}}
\newcommand{\elicembed}{\mathrm{elic}_\mathrm{embed}}

@@

\begin{proposition}\label{prop:embed-risk-poly}
  A loss $L$ embeds a discrete loss if and only if $\risk{L}$ is polyhedral.
  \jessie{Ready for a look over.}
\end{proposition}
\begin{proof}
  For the forward direction, suppose $L$ embeds a discrete loss $\ell$.
  By Proposition~\ref{prop:embed-bayes-risks}, we know that $\risk{L} = \risk{\ell}$.
  Since $\ell$ is discrete, we know that $\risk{\ell}$ is polyhedral as $\risk{\ell}$ is the minimum of a finite set of linear functions in $\simplex$. %\raf{do we know? okay, we do---but why?}.
  Therefore, $\risk{L}$ is polyhedral.

@@

>>>>

\newcommand{\reals}{\mathbb{R}}
\newcommand{\posreals}{\reals_{>0}}%{\reals_{++}}
\newcommand{\dom}{\mathrm{dom}}
\newcommand{\epi}{\text{epi}}
\newcommand{\prop}[1]{\Gamma[#1]}
\newcommand{\eliccts}{\mathrm{elic}_\mathrm{cts}}
\newcommand{\eliccvx}{\mathrm{elic}_\mathrm{cvx}}
\newcommand{\elicpoly}{\mathrm{elic}_\mathrm{pcvx}}
\newcommand{\elicembed}{\mathrm{elic}_\mathrm{embed}}

@@

\begin{proposition}\label{prop:embed-risk-poly}
  A loss $L$ embeds a discrete loss if and only if $\risk{L}$ is polyhedral.
\end{proposition}
\begin{proof}
  For the forward direction, suppose $L$ embeds a discrete loss $\ell$.
  By Proposition~\ref{prop:embed-bayes-risks}, we know that $\risk{L} = \risk{\ell}$.
  Since $\ell$ is discrete, we know that $\risk{\ell}$ is polyhedral as $\risk{\ell}$ is the minimum of a finite set of linear functions in $\simplex$. %\raf{do we know? okay, we do---but why?}.
  Therefore, $\risk{L}$ is polyhedral.

@@

\section{Top-$k$ surrogate}
Throughout this section, consider the surrogate and discrete loss $L'(u)_y~=~\left(\frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+$ and $\ell'(r)_y = \Ind{y \not\in r} + \frac{1}{k} |r \setminus \{y\}|$ given in Equations~\ref{eq:L-2-surrogate} and~\ref{eq:ell-2}, respectively.

\begin{lemma}\label{lem:top-k-polyhedral}
	$L'$ is a polyhedral loss.
\end{lemma}
\begin{proof}
	Observe $L'$ can be written as the pointwise max of $\binom{n}{k} +1$ terms, where the $\binom{n}{k}$ terms are selecting the $k$ elements of $u + \ones - e_y$, and the max comes from selecting the $u_i$ elements with highest weight.
\end{proof}
%By Lemma~\ref{lem:poly-loss-poly-risk}, we then know that $\risk{L'}$ is also polyhedral.

We next show that $\U = \{u \in \{0,1\}^n : \|u\|_0 \leq k\}$ is always represented among the optimizers of $L'$. \begin{lemma}\label{lem:top-k-optimal-corners}
	For all $p \in \simplex$ we have $\prop{L'}(p) \cap \U \neq \emptyset$.
\end{lemma}
\begin{proof}
	Let us first simplify $L'$:
	\[
	L'(u)_y = \left( 1 - u_y + \frac{1}{k} \sum_{i=1}^k (u - e_y)_{[i]} \right)_+~.
	\]
	Examining this expression, observe that
	if $u_i < u_{[k+1]}$, then $(u-e_y)_i < (u-e_y)_{[k]}$, as only one entry changes (and decreases) from one to the other.
	Hence, if $u_i < u_{[k+1]}$, then the index $i$ will not appear in the summation for any $y$, and thus only influences the loss when $y=i$.
	Now consider the loss of $u'$ which is the same as $u$ but with $u_i' = u_{[k+1]}$.
	Then $L'(u')_y \leq L'(u)_y$ when $y=i$ and $L'(u')_y = L'(u)_y$ when $y\neq i$.
	From this we see that without loss of generality, $u_{[k+1]} = \cdots = u_{[n]}$.

	Next, we show that $L'$ is invariant in the $\ones$ direction:
	\begin{align*}
	L'(u + \alpha \ones)_y &= \left( 1 - (u_y + \alpha) + \frac{1}{k} \sum_{i=1}^k ((u + \alpha) - e_y)_{[i]} \right)_+\\
	&= \left( 1 - (u_y + \alpha) + \frac{1}{k}\left[ \alpha k + \sum_{i=1}^k (u - e_y)_{[i]} \right] \right)_+\\
	&= \left( 1 -u_y + \frac{1}{k}\left[ \sum_{i=1}^k (u - e_y)_{[i]} \right] \right)_+\\
	&= L'(u)_y~.
	\end{align*}
	Thus, without loss of generality, we now have $u_{[k+1]} = \cdots = u_{[n]} = 0$.

	Let $u_{[1..j]} := \sum_{i=1}^j u_{[i]}$.
	We now claim that for all $y \in \Y$, the minimum of $L(u)_y$ is achieved by some $u \in [0,1]^n$.
	\raf{Only hole left is the following line.  I think it might not be true.}
	This follows from the observation that if $u_i > 1$, the the loss only decreases by reporting $u'$ which is equal to $u$ but with $u'_i = 1$.
	For $u\in[0,1]^n$, the argument to $(\cdot)_+$ is nonnegative, so we can rewrite the expected loss of $L'$ as
	\begin{align*}
	\inprod{p}{L'(u)} &= 1 - \inprod{p}{u} + \frac{1}{k} \left( \sum_{i=1}^k (1 - p_{[i]}) u_{[i]} \right) = 1 + \inprod{\tfrac 1 k \ones - (1+\tfrac 1 k)p}{u}~.
	\end{align*}
	As a linear objective over the hypercube, $\inprod{p}{L'}$ must be optimal on some corner of the hypercube, so we conclude $u_{[1]}, u_{[2]}, \ldots, u_{[k]} \in \{0,1\}$.

	In particular, for all $p \in \simplex$, there is a report $u \in \{0,1\}^n$ optimizing $L'$; since we can take $\|u\|_0 \leq k$ by the above observation that $u_{[k+1]} = u_{[k+2]} = \ldots = u_{[n]} = 0$, some $u \in \U$ must be optimal as well.
\end{proof}

\raf{Cut this final lemma for now}
% \begin{lemma}\label{lem:top-k-surrogate-embeds}
% $L':\reals^n \to \reals^\Y_+$ embeds $\ell':\R \to \reals^\Y_+$.
% \end{lemma}
% \begin{proof}
%   We can verify $\ell'(r) = L'(\varphi(r))$ for all $r \subseteq [n]$ and $\varphi(r)$ being its conversion from a set to binary vector, so $\varphi(\R) = \{ u \in \{0,1\}^n : \|u\|_1 \leq k \}$.
%   Moreover, by Lemma~\ref{lem:top-k-optimal-corners} and the above equality, respectively, we know that $\inf_{u \in \reals^n} \inprod{p}{L'(u)} = \min_{u \in \varphi(\R)} \inprod{p}{L'(u)} = \min_{r \in \R}\inprod{p}{\ell'(r)}$ for all $p \in \simplex$, and so we conclude that $L'$ embeds $\ell'$.
% \end{proof}

\hrule
\begin{proof}
	Let us first simplify $L'$:
	\[
	L'(u)_y = \left( 1 - u_y + \frac{1}{k} \sum_{i=1}^k (u - e_y)_{[i]} \right)_+~.
	\]
	Examining this expression, observe that
	if $u_i < u_{[k+1]}$, then $(u-e_y)_i < (u-e_y)_{[k]}$, as only one entry changes (and decreases) from one to the other.
	Hence, if $u_i < u_{[k+1]}$, then the index $i$ will not appear in the summation for any $y$, and thus only influences the loss when $y=i$.
	Now consider the loss of $u'$ which is the same as $u$ but with $u_i' = u_{[k+1]}$.
	Then $L'(u')_y \leq L'(u)_y$ when $y=i$ and $L'(u')_y = L'(u)_y$ when $y\neq i$.
	From this we see that without loss of generality, $u_{[k+1]} = \cdots = u_{[n]}$.

	Next, we show that $L'$ is invariant in the $\ones$ direction:
	\begin{align*}
	L'(u + \alpha \ones)_y &= \left( 1 - (u_y + \alpha) + \frac{1}{k} \sum_{i=1}^k ((u + \alpha) - e_y)_{[i]} \right)_+\\
	&= \left( 1 - (u_y + \alpha) + \frac{1}{k}\left[ \alpha k + \sum_{i=1}^k (u - e_y)_{[i]} \right] \right)_+\\
	&= \left( 1 -u_y + \frac{1}{k}\left[ \sum_{i=1}^k (u - e_y)_{[i]} \right] \right)_+\\
	&= L'(u)_y~.
	\end{align*}
	%Thus, without loss of generality, we now have $u_{[k+1]} = \cdots = u_{[n]} = 0$.

	Let $u_{[1..j]} := \sum_{i=1}^j u_{[i]}$.
	We now claim that for all $y \in \Y$, the minimum of $L(u)_y$ is achieved by some $u \in [0,1]^n$.

	First, we set $u_{[1]} = 1$ by invariance in the $\ones$ direction.
	Given this, we observe $L'(u)_y = \left(\tfrac 1 k (u-e_y)_{[1,\ldots,k]}\right)_+$.
	\jessie{Want to claim that $u_{[k+1]} \geq 0$.}
	%  In this case, we want to claim that having $u_{[k+2]} \geq u_{[k+1]}$, since $u_{[k+2]}, \ldots, u_{[n]}$ do not affect the loss, unless corresponding to $y$, so we want to increase these values as much as possible without them being part of the summand. 
	(In this case, for $u_y <0$, increasing $u_y$ to $0$ would decrease loss, since it would not affect the summand.)
	Additionally, if $u_{[k+1]} <0$, then the average $\tfrac 1 k (u - e_y)_{[1, \ldots, k]} = \tfrac 1 k u_{[2,\ldots,k]}$, so setting $u_{[k+1]} = 0$ can only decrease the loss.  (By the same argument used for $u_{[k+2]}$.)
	Therefore, we can set $u_{[k+1]} \geq 0$.\jessie{Is this true?  Revisit}

	Since $u_{[n]} = u_{[k+1]} \geq 0$, we have that there is some $u \in [0,1]^n$ minimizing $L'$.

	For $u\in[0,1]^n$, the argument to $(\cdot)_+$ is nonnegative, so we can rewrite the expected loss of $L'$ as
	\begin{align*}
	\inprod{p}{L'(u)} &= 1 - \inprod{p}{u} + \frac{1}{k} \left( \sum_{i=1}^k (1 - p_{[i]}) u_{[i]} \right) = 1 + \inprod{\tfrac 1 k \ones - (1+\tfrac 1 k)p}{u}~.
	\end{align*}
	As a linear objective over the hypercube, $\inprod{p}{L'}$ must be optimal on some corner of the hypercube, so we conclude for all $p \in \simplex$, there is a report $u \in \{0,1\}^n$ optimizing $L'$.

	So let us ask when it is better to set $u_i$ to $0$: \jessie{similar argument to the body of paper}

	\begin{align*}
	\inprod{p}{L'(u) - L'(\bar u)} &= 1 + \inprod{\tfrac 1 k \ones - (1 + \tfrac 1 k) p}{u} - (1 + \inprod{\tfrac 1 k \ones - (1 + \tfrac 1 k) p}{\bar u})\\
	&= \tfrac 1 k - (1+ \tfrac 1 k p_i) u_i - \tfrac 1 k - (1+ \tfrac 1 k p_i) \bar u_i \\
	&= \tfrac 1 k - (1 + \tfrac 1 k) p_i
	\end{align*}
	We can then see this term is positive only if $p_i > \tfrac 1 {k+1}$.

	Since at most $k$ elements can simultaneously have weight greater than $k+1$, we can then take $u_{[k+1]} = u_{[k+2]} = \ldots = u_{[n]} = 0$.
	We then observe an optimal report in $\U$ for all $p \in \simplex$.
	%  (On the uniform distribution, we argue that $\vec 0$ is also an optimal report, and so there is a report in $\U$ optimizing $L'$.)
	%  In particular, for all $p \in \simplex$, there is a report $u \in \{0,1\}^n$ optimizing $L'$; since we can take $\|u\|_0 \leq k$ by the above observation that $u_{[k+1]} = u_{[k+2]} = \ldots = u_{[n]} = 0$, some $u \in \U$ must be optimal as well.
\end{proof}
\hrule
\begin{lemma}
	For all $p \in \simplex$, there is a $u \in [0,1]^n$ optimizing $L'$.
\end{lemma}
\begin{proof}
	First, observe that for all $y \in \Y$, we have $L'(\cdot)_y$ bounded from below, so $\inprod p{L'(\cdot)}$ is also bounded from below for all $p \in \simplex$.

	Observe that for each point mass $\delta_y$, the expected loss $\inprod{\delta_y}{L'(\cdot)} = L'(\cdot)_y$ can be minimized by reporting $e_y$.
	Since every $p \in \simplex$ is a convex combination of point mass distributions, we claim the expected loss can be minmized by reporting some $u \in [0,1]^n$: the Minkowski sum of the $e_y$ vectors.

	Note that the Minkowski sum of the $e_y$ vectors is $[0,1]^n$ and is a subset of the full set of minimizers of $L'$ over $\simplex$.

	\jessie{Another thought: Sufficient to show $u_{[k+1]} \geq u_{[1]} - 1$?}

	Call $f_i(u) := L'(u)_i|_{[0,1]^n}$ the loss on outcome $i$ restricted to $[0,1]^n$.
	We know that all $f_i$ are polyhedral, since they are simply a (convex) restriction of polyhedral losses.
	By Rockafellar~\cite[Theorem 19.2]{rockafellar1997convex}, we then know $f^*_i$ is also polyhedral.
	Moreover, we know that $[0,1]^n = \dom f^*_i$ for all $i$ since $f_i$ is bounded from below, and so the intersection $\dom f^*_1 \cap \ldots \cap \dom f^*_n = [0,1]^n$ is nonempty.
	We can then apply Rockafellar~\cite[Corollary 20.1.1]{rockafellar1997convex} to observe the infimal convolution $f_1 \Box \ldots \Box f_n$ is polyhedral and its infimum is always attained.
	Since this term is a polyhedral function, its epigraph is a polyhedral set by definition.
	We know that $\epi(f_1 \Box \ldots \Box f_n) = \epi(f_1) + \epi(f_2) + \ldots + \epi(f_n)$, where $+$ is the Minkowski sum. \jessiet{cite... current source is Wikipedia}
	Moreover, $\epi(f_1) + \ldots \epi(f_n)$ is defined on $[0,1]^n$ and the infimum of the function sum is attained, since the epigraph sum is bounded from below  \jessie{not sure if that's what I want to say, but something like it.} on the function value.

	We know the infimum attained by taking a convex combination over $\{f_y\}_{y \in \Y}$ is the same as any over $\{L'_y\}_{y \in \Y}$, since, for all $y \in \Y$, we have $\inf_{u \in \reals^n} L'(u)_y = \inf_{u \in [0,1]^n} L'(u)_y$, so we cannot strictly improve expected loss by changing the minimizing report on any outcome.\jessie{??}
\end{proof}
